---
title: "COVID-19 Test"
output:
  # html_document:
  #   df_print: paged
  # html_notebook: default
---
### Load Libraries
```{r}
library(tidyverse)
library(tidycensus)
library(ggrepel)
library(tigris)
library(leaflet)
library(maps)
library(ggmap)
library(gganimate)
library(gifski)
library(transformr)
library(magick)
```

### Read in the NY Times Data
```{r}
# Get the NY Times data from github
countiesURL <- "https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv"
countiesData <- read_csv(url(countiesURL))
# Fix NYC
countiesData[which(countiesData$county == "New York City" & countiesData$state == "New York"),"fips"] <- "NYC"
# Fix KC Mo
kcCounties = c("Cass", "Clay", "Jackson", "Platte", "Kansas City")
kcCountyData = countiesData[which(countiesData$county %in% kcCounties & countiesData$state == "Missouri"),]

kcCountyData = kcCountyData %>%
  select(-c(county,state,fips))

kcCountyData <- kcCountyData %>%
  group_by(date) %>%
  summarise_all(sum)

kcCountyData = tibble(date = kcCountyData$date, county = "Kansas City", state = "Missouri", fips = "KC", cases = kcCountyData$cases, deaths = kcCountyData$deaths)
countiesData = countiesData[-which(countiesData$county %in% kcCounties & countiesData$state == "Missouri"),]
countiesData = rbind(countiesData,kcCountyData)
```

### Use Census API to get the populations of each county
```{r}
# Get census API key from: https://api.census.gov/data/key_signup.html
# census_api_key("YOUR KEY GOES HERE")
census_api_key("779c8b33af41fa4dbf19a22405c65c780fd379ac")

# Get population estimates per county from the US Census API
popEst <- get_estimates(geography = "county",
                        product = "population")
popEst <- popEst %>%
  filter(variable == "POP")

# Fix NYC
nycCounties = c("New York County, New York", "Kings County, New York", "Queens County, New York", 
                "Bronx County, New York", "Richmond County, New York")
nycCounts = sum(popEst[which(popEst$NAME %in% nycCounties),"value"])
nycCounts = tibble(NAME = "New York City, New York", GEOID = "NYC", variable = "POP", value = nycCounts)
popEst = popEst[-which(popEst$NAME %in% nycCounties),]
popEst = rbind(popEst,nycCounts)

# Fix KC Mo -- here, combine Cass, Clay, Jackson and Platte with KC
kcCounties = c("Cass County, Missouri", "Clay County, Missouri", "Jackson County, Missouri", "Platte County, Missouri")
kcCounts = sum(popEst[which(popEst$NAME %in% kcCounties),"value"])
kcCounts = tibble(NAME = "Kansas City, Missouri", GEOID = "KC", variable = "POP", value = kcCounts)
popEst = popEst[-which(popEst$NAME %in% kcCounties),]
popEst = rbind(popEst,kcCounts)
```

### Join the census data with the NY Times Data
# Need to fix places like NYC, etc.
```{r}
# Join Data
countiesData <- left_join(countiesData, popEst, by = c("fips" = "GEOID"))

countiesData <- countiesData %>%
  filter(!is.na(value)) # Here we are removing all the "Unknown" -- need to update this.... potentially use some of Paul's suggestions?

# Get cases and deaths per million population
countiesData <- countiesData %>%
  mutate(casesPerMillion = (cases/value)*1000000) %>%
  mutate(deathsPerMillion = (deaths/value)*1000000) 
```

### Set t = 0 to the first observed case in each county
```{r}
# Set t=0 to the date of the first case
time_zero <- countiesData %>%
  group_by(state, county) %>%
  summarise(first_case = min(date)) %>%
  ungroup

# Set a new column for the time elapsed between the date column and the t=0 date for each row
countiesData <- countiesData %>%
  left_join(time_zero, by = c("state", "county")) %>%
  mutate(time = as.numeric(date - first_case))

# Add a label for the max time for each county
max_date_label <- countiesData %>%
  group_by(state, county) %>%
  summarise(max_time = max(time))

countiesData <- countiesData %>%
  left_join(max_date_label, by = c("state", "county"))
```

### Generate plots for some case studies (can easily scale these up)
```{r}
#####################
# Raw counts
#####################
# Parameters
stateUse = "California"
minCases = 25
startDate = "2020-03-06"
endDate = "2020-03-27"
# Plot

selected_counties <- countiesData %>%
  group_by(state, county) %>%
  summarise(max_cases_per_county = max(cases)) %>%
  mutate(has_enough_cases = (max_cases_per_county > minCases)) %>%
  filter(has_enough_cases) %>%
  ungroup

countiesData <- countiesData %>%
  left_join(selected_counties, by = c("state", "county"))
```

### Linear Scale
```{r fig.height=6, fig.width=8}
countiesData %>%
  filter(state == "California" & 
           date >= startDate &
           date <= endDate &
           has_enough_cases) %>%
    mutate(label = if_else(date == max(date), as.character(county), NA_character_)) %>%
  ggplot(aes(x = date, y = cases, group = county, color = county)) + 
  geom_line() + 
  geom_point() +
  geom_label_repel(aes(label = label),
                   nudge_x = 1,
                   na.rm = TRUE) +
  ggtitle(label = paste0("Cumulative Cases in Each ", stateUse, " County - Linear Plot"), subtitle = paste0("minimum each county is ",minCases," cases")) +
  xlab(paste0("Date starting: ",startDate)) +
  ylab("Cumulative Cases") +
  theme_bw(base_size = 12) +
  theme(legend.position = "none")
```

### Log Scale
```{r fig.height=6, fig.width=8}
# Log scale
countiesData %>%
  filter(state == "California" & 
           date >= startDate &
           date <= endDate &
           has_enough_cases) %>%
  mutate(label = if_else(date == max(date), as.character(county), NA_character_)) %>%
  ggplot(aes(x = date, y = cases, group = county, color = county)) + 
  geom_line() + 
  geom_point() +
  geom_label_repel(aes(label = label),
                   nudge_x = 1,
                   na.rm = TRUE) +
  scale_y_continuous(trans='log10') +
  ggtitle(label = paste0("Cumulative Cases in Each ", stateUse, " County - Log Plot"), subtitle = paste0("minimum each county is ",minCases," cases")) +
  xlab(paste0("Date starting: ",startDate)) +
  ylab("Cumulative Cases") +
  theme_bw(base_size = 12) +
  theme(legend.position = "none")
```

### Log Scale -- cases per million
```{r fig.height=6, fig.width=8}
# Log scale
countiesData %>%
  filter(state == "California" & 
           date >= startDate &
           date <= endDate &
           has_enough_cases) %>%
  mutate(label = if_else(date == max(date), as.character(county), NA_character_)) %>%
  ggplot(aes(x = date, y = casesPerMillion, group = county, color = county)) + 
  geom_line() + 
  geom_point() +
  geom_label_repel(aes(label = label),
                   nudge_x = 1,
                   na.rm = TRUE) +
  scale_y_continuous(trans='log10') +
  ggtitle(label = paste0("Cumulative Cases Per Million in Each ", stateUse, " County - Log Plot"), subtitle = paste0("minimum each county is ",minCases," cases")) +
  xlab(paste0("Date starting: ",startDate)) +
  ylab("Cumulative Cases Per Million") +
  theme_bw(base_size = 12) +
  theme(legend.position = "none")
```

### Log Scale -- cases per million -- first case, t=0
```{r fig.height=6, fig.width=8}
# Log scale
countiesData %>%
  filter(state == "California" & 
           has_enough_cases) %>%
  mutate(label = if_else(time == max_time, as.character(county), NA_character_)) %>%
  ggplot(aes(x = time, y = casesPerMillion, group = county, color = county)) + 
  geom_line() + 
  geom_point() +
  geom_label_repel(aes(label = label),
                   nudge_x = 1,
                   na.rm = TRUE) +
  scale_y_continuous(trans='log10') +
  ggtitle(label = paste0("Cumulative Cases Per Million in Each ", stateUse, " County - Log Plot"), subtitle = paste0("minimum each county is ",minCases," cases")) +
  xlab("Time Since First Case") +
  ylab("Cumulative Cases Per Million") +
  theme_bw(base_size = 12) +
  theme(legend.position = "none")
```

### Generate an interactive map
```{r message=FALSE, warning=FALSE}
# This is based of off the tutorial here: http://zevross.com/blog/2015/10/14/manipulating-and-mapping-us-census-data-in-r-using-the-acs-tigris-and-leaflet-packages-3/

# grab the spatial data (tigris)
# Get the counties of interest
countiesUse <- unique(countiesData$county[which(countiesData$state=="California")])

# Map the counties to the spatial data
tracts <- tracts(state = "California", county = countiesUse, cb = TRUE)
tracts$fips = paste0(tracts$STATEFP,tracts$COUNTYFP)

# Join our data frame with the counts data to the spatial data
counts_merged <- geo_join(tracts, countiesData[which(countiesData$state=="California" & countiesData$date == endDate),], "fips", "fips", how = 'inner')
# there are some tracts with no land that we should exclude
counts_merged <- counts_merged[counts_merged$ALAND>0,]

# Setup the pop-up that comes up when hovering
popup <- paste0("County: ", counts_merged$county, "<br>", "Cases per million: ", round(counts_merged$casesPerMillion,2))
pal <- colorNumeric(
  palette = "YlGnBu",
  domain = counts_merged$casesPerMillion
)

# Plot everything
map3<-leaflet() %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(data = counts_merged, 
              fillColor = ~pal(casesPerMillion), 
              color = "#b2aeae", # you need to use hex colors
              fillOpacity = 0.7, 
              weight = 1, 
              smoothFactor = 0.2,
              popup = popup) %>%
  addLegend(pal = pal, 
            values = counts_merged$casesPerMillion, 
            position = "bottomright", 
            title = "Cases per million") 
map3
```

### Generate gif for just California
```{r}
# Get the boundaries of the states and counties
states <- map_data("state")
state_df <- subset(states, region == tolower("California"))
counties <- map_data("county")
state_county <- subset(counties, region == tolower("California"))

# Get the base map for the state
state_base <- ggplot(data = state_df, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
  geom_polygon(color = "black", fill = "gray")

# Join the state spatial data with the covid-19 data
# stateSubset = countiesData[which(countiesData$state=="California" & countiesData$date == endDate),]
stateSubset = countiesData[which(countiesData$state=="California"),]
stateSubset$county = tolower(stateSubset$county)
stateSubset <- inner_join(state_county, stateSubset, by = c("subregion" = "county"))

# By date
p1 <- state_base + 
  geom_polygon(data = stateSubset, aes(fill = casesPerMillion), color = "white") +
  geom_polygon(color = "black", fill = NA) +
  scale_fill_viridis_c(trans = "log2") +
  # Here comes the gganimate specific bits
  labs(title = 'Date: {frame_time}') +
  transition_time(date) +
  # ease_aes('linear') +
  # scale_fill_gradient(trans = "log10") +
  theme_bw() +
  theme(
    axis.text = element_blank(),
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    panel.border = element_blank(),
    panel.grid = element_blank(),
    axis.title = element_blank())
p1

anim_save(filename = "./CaliforniaCasesPerMillion.gif",animation = animate(p1, nframes = 100, duration = 40, fps = 10, end_pause = 20, rewind = FALSE))

# Slo-mo
p1_gif <- animate(p1, width = 480, height = 480, nframes = 100, duration = 10, fps = 10, end_pause = 10, rewind = FALSE)

p1_mgif <- image_read(p1_gif)

p1_new_gif <- image_append(c(p1_mgif[1]))
for(i in 2:100){
  if(i <= 50){
    combined <- image_append(c(p1_mgif[i]))
    p1_new_gif <- c(p1_new_gif, combined)
  } else{
    if(i > 50 & i <= 75){
      combined <- image_append(c(p1_mgif[i]))
      p1_new_gif <- c(p1_new_gif, combined)
      p1_new_gif <- c(p1_new_gif, combined)
      p1_new_gif <- c(p1_new_gif, combined)
      p1_new_gif <- c(p1_new_gif, combined)
      p1_new_gif <- c(p1_new_gif, combined)
    }
    else{
      combined <- image_append(c(p1_mgif[i]))
      p1_new_gif <- c(p1_new_gif, combined)
      p1_new_gif <- c(p1_new_gif, combined)
      p1_new_gif <- c(p1_new_gif, combined)
      p1_new_gif <- c(p1_new_gif, combined)
      p1_new_gif <- c(p1_new_gif, combined)
      p1_new_gif <- c(p1_new_gif, combined)
      p1_new_gif <- c(p1_new_gif, combined)
      p1_new_gif <- c(p1_new_gif, combined)
      p1_new_gif <- c(p1_new_gif, combined)
      p1_new_gif <- c(p1_new_gif, combined)
    }
  }
}

anim_save(filename = "./CaliforniaCasesPerMillionSloMo.gif",animation = p1_new_gif)


# TRY TIME SINCE FIRST CASE?!?!
# p1 <- state_base + 
#   geom_polygon(data = stateSubset, aes(fill = casesPerMillion), color = "white") +
#   geom_polygon(color = "black", fill = NA) +
#   # Here comes the gganimate specific bits
#   labs(title = 'Time since first case: {frame_time}') +
#   transition_time(time) +
#   # ease_aes('linear') +
#   # scale_fill_gradient(trans = "log10") +
#   theme_bw() +
#   theme(
#     axis.text = element_blank(),
#     axis.line = element_blank(),
#     axis.ticks = element_blank(),
#     panel.border = element_blank(),
#     panel.grid = element_blank(),
#     axis.title = element_blank())
# p1
```

### Generate gif for multiple states
```{r}

# Specify a new start date -- make the video better
newStart = as.Date("2020-03-04")

# Get the max value to make all the scales the same range
maxVal = max(countiesData$casesPerMillion[which(countiesData$state %in% c("California","Washington","Texas","New York"))])

# Get the boundaries of the states and counties
states <- map_data("state")
ca_df <- subset(states, region == tolower("California"))
wa_df <- subset(states, region == tolower("Washington"))
tx_df <- subset(states, region == tolower("Texas"))
ny_df <- subset(states, region == tolower("New York"))

counties <- map_data("county")
# Fix for NYC
nycCounties = c("new york","kings","queens","bronx","richmond")
counties[which(counties$region == "new york" & counties$subregion %in% nycCounties),"subregion"] <- "new york city"

ca_county <- subset(counties, region == tolower("California"))
wa_county <- subset(counties, region == tolower("Washington"))
tx_county <- subset(counties, region == tolower("Texas"))
ny_county <- subset(counties, region == tolower("New York"))

# Get the base map for the state
ca_base <- ggplot(data = ca_df, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
  geom_polygon(color = "black", fill = "gray")

wa_base <- ggplot(data = wa_df, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
  geom_polygon(color = "black", fill = "gray")

tx_base <- ggplot(data = tx_df, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
  geom_polygon(color = "black", fill = "gray")

ny_base <- ggplot(data = ny_df, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
  geom_polygon(color = "black", fill = "gray")

# Join the state spatial data with the covid-19 data
# stateSubset = countiesData[which(countiesData$state=="California" & countiesData$date == endDate),]
caSubset = countiesData[which(countiesData$state=="California"),]
caSubset$county = tolower(caSubset$county)
caSubset <- inner_join(ca_county, caSubset, by = c("subregion" = "county"))

waSubset = countiesData[which(countiesData$state=="Washington"),]
waSubset$county = tolower(waSubset$county)
waSubset <- inner_join(wa_county, waSubset, by = c("subregion" = "county"))

txSubset = countiesData[which(countiesData$state=="Texas"),]
txSubset$county = tolower(txSubset$county)
txSubset <- inner_join(tx_county, txSubset, by = c("subregion" = "county"))

nySubset = countiesData[which(countiesData$state=="New York"),]
nySubset$county = tolower(nySubset$county)
nySubset <- inner_join(ny_county, nySubset, by = c("subregion" = "county"))

# By date
p1 <- ca_base + 
  geom_polygon(data = caSubset[which(caSubset$date>=newStart),], aes(fill = casesPerMillion), color = "white") +
  geom_polygon(color = "black", fill = NA) +
  scale_fill_viridis_c(limits=c(1,maxVal),trans = "log2") +
  labs(title = 'Date: {frame_time}') +
  transition_time(date) +
  theme_bw() +
  theme(
    axis.text = element_blank(),
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    panel.border = element_blank(),
    panel.grid = element_blank(),
    axis.title = element_blank())

p2 <- wa_base + 
  geom_polygon(data = waSubset[which(waSubset$date>=newStart),], aes(fill = casesPerMillion), color = "white") +
  geom_polygon(color = "black", fill = NA) +
  scale_fill_viridis_c(limits=c(1,maxVal),trans = "log2") +
  labs(title = 'Date: {frame_time}') +
  transition_time(date) +
  theme_bw() +
  theme(
    axis.text = element_blank(),
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    panel.border = element_blank(),
    panel.grid = element_blank(),
    axis.title = element_blank())

p3 <- tx_base + 
  geom_polygon(data = txSubset[which(txSubset$date>=newStart),], aes(fill = casesPerMillion), color = "white") +
  geom_polygon(color = "black", fill = NA) +
  scale_fill_viridis_c(limits=c(1,maxVal),trans = "log2") +
  labs(title = 'Date: {frame_time}') +
  transition_time(date) +
  theme_bw() +
  theme(
    axis.text = element_blank(),
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    panel.border = element_blank(),
    panel.grid = element_blank(),
    axis.title = element_blank())

p4 <- ny_base + 
  geom_polygon(data = nySubset[which(nySubset$date>=newStart),], aes(fill = casesPerMillion), color = "white") +
  geom_polygon(color = "black", fill = NA) +
  scale_fill_viridis_c(limits=c(1,maxVal),trans = "log2") +
  labs(title = 'Date: {frame_time}') +
  transition_time(date) +
  theme_bw() +
  theme(
    axis.text = element_blank(),
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    panel.border = element_blank(),
    panel.grid = element_blank(),
    axis.title = element_blank())

a_gif <- animate(p1, width = 240, height = 240, nframes = 100, duration = 20, fps = 10, end_pause = 20, rewind = FALSE)
b_gif <- animate(p2, width = 240, height = 240, nframes = 100, duration = 20, fps = 10, end_pause = 20, rewind = FALSE)
c_gif <- animate(p3, width = 240, height = 240, nframes = 100, duration = 20, fps = 10, end_pause = 20, rewind = FALSE)
d_gif <- animate(p4, width = 240, height = 240, nframes = 100, duration = 20, fps = 10, end_pause = 20, rewind = FALSE)

a_mgif <- image_read(a_gif)
b_mgif <- image_read(b_gif)
c_mgif <- image_read(c_gif)
d_mgif <- image_read(d_gif)

new_gif <- image_append(c(a_mgif[1], b_mgif[1], c_mgif[1], d_mgif[1]))
for(i in 2:200){
  combined <- image_append(c(a_mgif[i], b_mgif[i], c_mgif[i], d_mgif[i]))
  new_gif <- c(new_gif, combined)
}

# new_gif

anim_save(filename = "./MultiStateCasesPerMillion.gif",animation = new_gif)



# Slo-mo
a_gif <- animate(p1, width = 240, height = 240, nframes = 100, duration = 10, fps = 10, end_pause = 5, rewind = FALSE)
b_gif <- animate(p2, width = 240, height = 240, nframes = 100, duration = 10, fps = 10, end_pause = 5, rewind = FALSE)
c_gif <- animate(p3, width = 240, height = 240, nframes = 100, duration = 10, fps = 10, end_pause = 5, rewind = FALSE)
d_gif <- animate(p4, width = 240, height = 240, nframes = 100, duration = 10, fps = 10, end_pause = 5, rewind = FALSE)

a_mgif <- image_read(a_gif)
b_mgif <- image_read(b_gif)
c_mgif <- image_read(c_gif)
d_mgif <- image_read(d_gif)

new_gif <- image_append(c(a_mgif[1], b_mgif[1], c_mgif[1], d_mgif[1]))
for(i in 2:100){
  if(i <= 50){
    combined <- image_append(c(a_mgif[i], b_mgif[i], c_mgif[i], d_mgif[i]))
    new_gif <- c(new_gif, combined)
  } else{
    if(i > 50 & i <= 75){
      combined <- image_append(c(a_mgif[i], b_mgif[i], c_mgif[i], d_mgif[i]))
      new_gif <- c(new_gif, combined)
      new_gif <- c(new_gif, combined)
      new_gif <- c(new_gif, combined)
      new_gif <- c(new_gif, combined)
      new_gif <- c(new_gif, combined)
    }
    else{
      combined <- image_append(c(a_mgif[i], b_mgif[i], c_mgif[i], d_mgif[i]))
      new_gif <- c(new_gif, combined)
      new_gif <- c(new_gif, combined)
      new_gif <- c(new_gif, combined)
      new_gif <- c(new_gif, combined)
      new_gif <- c(new_gif, combined)
      new_gif <- c(new_gif, combined)
      new_gif <- c(new_gif, combined)
      new_gif <- c(new_gif, combined)
      new_gif <- c(new_gif, combined)
      new_gif <- c(new_gif, combined)
    }
  }
}

anim_save(filename = "./MultiStateCasesPerMillionSloMo.gif",animation = new_gif)



```

### Try NYC and Mo plots
```{r}
# Get the boundaries of the states and counties
states <- map_data("state")
ny_df <- subset(states, region == tolower("New York"))

counties <- map_data("county")
# Fix for NYC
nycCounties = c("new york","kings","queens","bronx","richmond")
counties[which(counties$region == "new york" & counties$subregion %in% nycCounties),"subregion"] <- "new york city"
ny_county <- subset(counties, region == tolower("New York"))

# Get the base map for the state
ny_base <- ggplot(data = ny_df, mapping = aes(x = long, y = lat, group = group)) + 
  coord_fixed(1.3) + 
  geom_polygon(color = "black", fill = "gray")

# Join the state spatial data with the covid-19 data
nySubset = countiesData[which(countiesData$state=="New York"),]
nySubset$county = tolower(nySubset$county)
nySubset <- inner_join(ny_county, nySubset, by = c("subregion" = "county"))

p.ny <- ny_base + 
  geom_polygon(data = nySubset[which(nySubset$date=="2020-03-27"),], aes(fill = casesPerMillion), color = "white") +
  geom_polygon(color = "black", fill = NA) +
  scale_fill_viridis_c(trans = "log2")
p.ny

# By date
p.ny <- ny_base + 
  geom_polygon(data = nySubset, aes(fill = casesPerMillion), color = "white") +
  geom_polygon(color = "black", fill = NA) +
  scale_fill_viridis_c(trans = "log2") +
  labs(title = 'Date: {frame_time}') +
  transition_time(date) +
  theme_bw() +
  theme(
    axis.text = element_blank(),
    axis.line = element_blank(),
    axis.ticks = element_blank(),
    panel.border = element_blank(),
    panel.grid = element_blank(),
    axis.title = element_blank())

```







